A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-07-01 elsfrontmatter.tex
sjs Revisions

\author[pbl]{Maarten van den Berg}
\author[nies]{Shinichiro Fujimori}
\author[pik]{David Klein}
\author[iiasa]{Keywan Riahi}
\author[pnnl]{Steven Smith}
\author[pbl]{Detlef van Vuuren}

@@

\begin{abstract}
Harmonization describes the process required to compare Integrated Assessment
Model (IAM) GHG and air pollutants results against a common source of historical
emissions trajectories. To date, harmonization has been performed separately by
individual modeling teams through
% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
% 
% > harmonization has been performed separately by individual modeling teams 
% I think harmonization is not done. We just report the direct model outputs.
% 
% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
% 
% ^.
the use of expert opinion and solicitation. In order to support ongoing and
future global climate and socioeconomic analyses, incorporate model advancements
towards more complex regional and sectoral representations, and expand the
capability of modeling teams to participate in intercomparison exercises, an
automated approach to choosing harmonization methods becomes necessary. This
work describes a novel
% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
% 
% > novel
% It would be better to stress how novel and useful.
% 
% ^.
methodology for determining such harmonization methods and an open-source Python
code base for implementing the methodology. Results are shown for two example
scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
of the total emissions trajectories.
\end{abstract}

>>>>

\author[pbl]{Maarten van den Berg}
\author[nies]{Shinichiro Fujimori}
\author[pik]{David Klein}
\author[iiasa]{Keywan Riahi}
\author[pnnl]{Steven J. Smith}
\author[pbl]{Detlef van Vuuren}

@@

\begin{abstract}
Harmonization describes the process required to match Integrated Assessment
Model (IAM) GHG and air pollutants results against a common source of historical
emissions trajectories. To date, harmonization has been performed separately by
individual modeling teams through
% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
% 
% > harmonization has been performed separately by individual modeling teams 
% I think harmonization is not done. We just report the direct model outputs.
% 
% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
% 
% ^.
the use of expert opinion and solicitation. In order to support ongoing and
future global climate and socioeconomic analyses, incorporate model advancements
towards more complex regional and sectoral representations, and expand the
capability of modeling teams to participate in intercomparison exercises, an
automated approach to choosing harmonization methods becomes necessary. This
work describes a novel
% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
% 
% > novel
% It would be better to stress how novel and useful.
% 
% ^.
methodology for determining such harmonization methods and an open-source Python
code base for implementing the methodology. Results are shown for two example
scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
of the total emissions trajectories.
\end{abstract}

### 2017-07-01 intro.tex
sjs Revisions

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are primarily designed
to be used in analyses of broad climate research which include earth system
model (ESM) simulations, climate impact, adaptation and climate mitigation
studies. They are regularly assessed by the United Nations' Intergovernmental
Panel on Climate Change (IPCC) across working groups.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, outputs, and purpose are similar across different
models. Modeling teams incorporate existing data sets of energy systems, land
use, economics, demographics and emissions sources and concentrations, among other
data, in order to provide consistent historical trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies (e.g., NOx and SOx scrubbers on power plants). However, the
historical emissions inventories calculated by different models can differ by
large amounts depending on the region, sector, and emissions species being
modeled.

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plants in China, and incorporate
emissions-intensity factors for individual gas species, for example the quantity
of sulfur emissions from coal plants per megawatt-hour of production, in order
to determine total emissions of a given gas species from a given
technology. Models are then \textit{calibrated} to historical data sources which
provide a combination of the activity of emitting technologies and trajectories
of past emissions in order to determine historical emissions trajectories in
each model region for each technology. Due to the sometimes large uncertainties
in historical datasets, results in the historical period may differ between
models.

With some frequency, historical data is updated to incorporate new assumptions
of source activity, new emissions accounting practices, or other concerns. The
global climate change community has recently developed a standard historical
emissions dataset for both anthropogenic emissions (i.e., the Community
Emissions Data System (CEDS) \cite{hoesly_historical_2017} and land-use change
(LUC) emissions \cite{van_marle_historic_2017}) which, in
% * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
% 
% > land-use change (LUC) emissions \cite{van_marle_historic_2017}
% I think this one also anthropogenic
% 
% ^.
conjunction with the SSP IAM trajectories, will be used for climate-related
modeling exercises of CMIP6.

When participating in intercomparison exercises in which consistency of
historical trajectories is required (e.g., in CMIP6), model teams incorporate a
single, common historical dataset through \textit{harmonization} in order to
enhance comparability across models. Harmonization describes the process of
adjusting model results to incorporate new historical time series while
maintaining critical aspects of and providing a smooth transition to modeled
future trajectories such that the resulting trajectories are consistent with the
original modeled results and usable by other members of the scientific community
involved in the exercise. In the emissions context, this means that each
individual combination of model region, model sector, and emissions species must
be harmonized. Depending on the total number of model regions, sectors, and
emissions species, this can require the selection of thousands to
tens-of-thousands of harmonization methods.

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution. IAMs provide global projections of
systemic change by dividing the world into a number of representative regions
(typically 10 to 30), the definition of which is distinct for each model
\cite{krey_global_2014}. Results from IAMs are integral in a number of
international studies, which notably include projections of climate and economic
futures. Recently, flagship IAMs have developed scenarios based on the Shared
% * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
% 
% > flagship
% Suggest different word. "comprehensive" is one possibility.
% 
% ^ <ssmith@pnnl.gov> 2017-07-01T03:11:57.571Z.
Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
quantify a variety of potential global futures. The SSPs are  designed
to be used in  research that include earth system
model (ESM) simulations, climate impact, adaptation and climate mitigation
studies.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, and outputs  are similar across different
models. Modeling teams incorporate  data on energy systems, land
use, economics, demographics and emissions sources and concentrations, among other
data, in order to provide consistent future trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies (e.g., NOx and SOx scrubbers on power plants). However, the
historical emissions starting points of  models can differ by
large amounts depending on the region, sector, and emissions species.

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plants in China, and incorporate
emissions-intensity factors for individual gas species, for example the quantity
of sulfur emissions from coal plants per megawatt-hour of production. Models are generally \textit{calibrated} to historical data sources in one or more base years. Due to the sometimes large uncertainties
in historical datasets, results in the historical period may differ between
models. Models also differ in their choice of base-year, which may lag behind available inventory data. In addition, models have varying sectoral, regional, and fuel aggregations.

The
global climate change community has recently developed a new global historical
emissions dataset for both anthropogenic  emissions (i.e., the Community
Emissions Data System (CEDS) \cite{hoesly_historical_2017} and open-burning land-use and land-use change (LULUC)
emissions \cite{van_marle_historic_2017}) which, in
% * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
% 
% > land-use change (LUC) emissions \cite{van_marle_historic_2017}
% I think this one also anthropogenic
% 
% ^.
conjunction with the SSP IAM trajectories, will be used for climate-related
modeling exercises of CMIP6.

When participating in intercomparison exercises in which a consistent historical starting point is required (e.g., in CMIP6), model teams incorporate a
single, common historical dataset through \textit{harmonization}. Harmonization describes the process of
adjusting model results to match the selected historical time series  such that the resulting future trajectories are also consistent with the
original modeled results. In the emissions context, this means that each
individual combination of model region, model sector, and emissions species must
be harmonized. Depending on the total number of model regions, sectors, and
emissions species, this can require the selection of thousands to
tens-of-thousands of harmonization methods.

### 2017-07-01 meth.tex
sjs Revisions

IAM emission results are provided along temporal (normally half decade or
decade), spatial (i.e., model regions), gas species, and sectoral
dimensions. Each individual temporal trajectory, i.e., unique combinations of
regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
initial time step. Given a model trajectory, $m_{r, g, s}(t)$, historical
trajectory, $h_{r, g, s}(t)$, and initial time step $t_i$, a harmonized
trajectory can be calculated. The ``goodness of fit'' of a trajectory depends on
a number of factors, most notably the faithful representation of the original,
unharmonized, trajectory. However, other important factors must be considered,
such as the representation of negative trajectories (i.e., if a trajectory
becomes negative, both the timing and total magnitude should be as close as
possible) which are of critical importance for cumulative CO2 calculations.

@@

Variability of the historical trajectory is also an important characteristic
when considering the choice of harmonization method. Such trajectories are
normally indicative of emissions from land use sectors. Land use sectors are
modeled with varying fidelity by IAMs, but in general proxies are used for
emissions trajectories. Take for example emissions from grassland burning: for
most models, the use of grass and paturelands is endogenously modeled, and
emissions from the burning of these lands is determined by applying an emissions
factor; in general, the more grassland usage that occurs in a scenario, the
higher the emissions from the burning thereof. Therefore, consistency in
harmonization method is important because the effects are modeled similarly
across regions and species. In order to reliably detect land-use-like emissions,
a measure of the coefficient of variation of the first derivative of the
historical trajectory, $c_v$, is calculated as shown in Equation \ref{eqs:cov},
where $\sigma$ is the standard deviation, and $\mu$ is the mean of the
trajectory.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{dh.pdf}
    \caption[]{
      \label{fig:dh}
      The distribution of relative differences between model and historical
      values in the harmonization year is shown. The solid black line indicates
      the threshold value used by default in \code{aneris}.  
    }
  \end{center}
\end{figure}

>>>>

IAM emission results are provided along temporal (normally half decade or
decade), spatial (i.e., model regions), gas species, and sectoral
dimensions. Each individual temporal trajectory, i.e., unique combinations of
regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
initial starting point. Given a model trajectory, $m_{r, g, s}(t)$, historical
values, $h_{r, g, s}(t)$, and initial starting point $t_i$, a harmonized
trajectory needs to be calculated. The ``goodness of fit'' of a trajectory depends on
a number of factors, most notably the faithful representation of the original,
unharmonized, trajectory. However, other important factors must be considered,
such as the representation of negative trajectories (i.e., if a trajectory
becomes negative, both the timing and total magnitude should be as close as
possible) which are of critical importance for cumulative CO2 calculations.
% * <ssmith@pnnl.gov> 2017-07-01T03:25:04.357Z:
% 
% Suggest moving this later where negative emissions are discussed. Its a bit of a tangent here.
% 
% ^ <ssmith@pnnl.gov> 2017-07-01T03:25:44.627Z.

@@

Temporal variability of the historical trajectory is also an important characteristic
when considering the choice of harmonization method.  Emissions from forest and grassland fires, for example, vary from year to year due to a combination of meteorological conditions and anthropogenic drivers. Land use emissions are modeled by IAMs using average emission factors and do not capture conditions in a specific year.  Ratio methods for such sectors can lead to unrealistic harmonization results.
modeled with varying fidelity by IAMs, but in general proxies are used for
emissions trajectories. Take for example emissions from grassland burning: for
most models, the use of grass and paturelands is endogenously modeled, and
emissions from the burning of these lands is determined by applying an emissions
factor; in general, the more grassland usage that occurs in a scenario, the
higher the emissions from the burning thereof. Therefore, consistency in
harmonization method is important because the effects are modeled similarly
across regions and species. In order to detect land-use-like emissions,
% * <ssmith@pnnl.gov> 2017-07-01T03:31:52.910Z:
% 
% > modeled with varying fidelity by IAMs, but in general proxies are used for
% > emissions trajectories. Take for example emissions from grassland burning: for
% > most models, the use of grass and paturelands is endogenously modeled, and
% > emissions from the burning of these lands is determined by applying an emissions
% > factor; in general, the more grassland usage that occurs in a scenario, the
% > higher the emissions from the burning thereof. Therefore, consistency in
% > harmonization method is important because the effects are modeled similarly
% > across regions and species. 
% 
% Not sure I follow the logic here. I added some new text that might help. Perhaps the intermediate part could be deleted.
% 
% ^ <ssmith@pnnl.gov> 2017-07-01T03:32:30.453Z.
a measure of the coefficient of variation of the first derivative of the
historical trajectory, $c_v$, is calculated as shown in Equation \ref{eqs:cov},
where $\sigma$ is the standard deviation, and $\mu$ is the mean of the
trajectory.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{dh.pdf}
    \caption[]{
      \label{fig:dh}
      The distribution of relative differences between model and historical
      values in the harmonization year is shown. The solid black line indicates
      the 50\% threshold value used by default in \code{aneris}.  
    }
  \end{center}
\end{figure}

### 2017-07-01 results.tex
sjs Revisions

In order to test the \code{aneris} harmonization procedure, results from the IAM
MESSAGE-GLOBIOM \cite{fricko_marker_2017} was processed. Two scenarios from the
SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
or ``middle of the road'', scenario family is chosen to be analyzed because
MESSAGE-GLOBIOM is the marker scenario for this SSP. Of all SSP2 scenarios, the
reference (marker) and SSP2-45 scenarios are analyzed. Notably, emissions from
multiple sectors tend to increase in the reference scenario, thus testing the
harmonization mechanism with an increasing emissions pathway. The SSP2-45
scenario, where the ``45'' designation refers to a scenario with end-of-century
emissions resulting in approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$
radiative forcing due to both warming from GHGs and fluoridated gases as well as
cooling from aerosols, is chosen because mitigation technologies and policies
are enacted causing a general reduction in pollutants and GHGs, including
(eventual) negative CO2 emissions in some regions and sectors due to carbon
capture and sequestration and afforestation. This scenario explores the
harmonization mechanism's response to generally decreasing emission
trajectories. The emissions of Kyoto Gases, shown in Figure \ref{fig:kyoto},
display the relative difference in trends between scenarios.

@@

The effect of harmonization overrides is analyzed by harmonizing each scenario
first without and then with overrides. Of the 1430 trajectories, approximately
10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 2\%
required the use of harmonization overrides after an initial investigation;
thus, 98\% of all trajectories were satisfactorily determined using the default
methods as chosen by the harmonization algorithm. Figure \ref{fig:nox} provides
one such example of a emissions species and sector in which all regions were
satisfactorily harmonized with the default methods. As can be seen, results with
and without overrides are identical.

@@

The harmonization of emissions pathways is performed in order to accurately
represent an updated historical emissions trajectory while also maintaining
consistency with the original, unharmonized pathway. As such, when the default
methods as provided by the harmonization procedure distort or otherwise
sufficiently misrepresent the underlying unharmonized results, an override
method is required to be provided for the trajectory of the region, sector, and
species in question. Based on the 2\% of total trajectories required to be
overridden, two classifications were observed: regional trajectories whose
\textit{magnitude} was overly distorted and regional trajectories whose
\textit{shape} was overly distorted.

>>>>

In order to test the \code{aneris} harmonization procedure, results from the IAM
MESSAGE-GLOBIOM \cite{fricko_marker_2017} was processed. Two scenarios from the
SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
or ``middle of the road'', scenario family is chosen to be analyzed because
MESSAGE-GLOBIOM is the marker scenario for this SSP. Of all SSP2 scenarios, the
reference (marker) and SSP2-45 scenarios are analyzed. Notably, emissions from
multiple sectors tend to increase in the reference scenario, thus testing the
harmonization mechanism with an increasing emissions pathway. The SSP2-45
scenario, where the ``45'' designation refers to a scenario with end-of-century
emissions resulting in approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$
radiative forcing due to both warming from GHGs and fluoridated gases as well as
cooling from aerosols, is chosen because mitigation technologies and policies
are enacted causing a general reduction in pollutants and GHGs, including
(eventual) negative CO2 emissions in some regions and sectors due to carbon
capture and sequestration and afforestation. This scenario explores the
harmonization mechanism's response to generally decreasing emission
trajectories. The emissions of Kyoto Gases, shown in Figure \ref{fig:kyoto},
shows the different trends of these two scenarios.

@@

The effect of harmonization overrides is analyzed by harmonizing each scenario
first without and then with overrides. Of the 1430 trajectories, approximately
10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 2\%
required the use of harmonization overrides after an initial investigation;
thus, 98\% of all trajectories were satisfactorily harmonized using the default
methods. Figure \ref{fig:nox} provides
one such example of a emissions species and sector in which all regions were
satisfactorily harmonized with the default methods. As can be seen, results with
and without overrides are identical.

@@

The harmonization of emissions pathways is performed in order to accurately
represent an updated historical emissions trajectory while also maintaining
consistency with the original, unharmonized pathway. As such, when the default
methods as provided by the harmonization procedure distort or otherwise
sufficiently misrepresent the underlying unharmonized results, an override
method is required to be provided for the trajectory of the region, sector, and
species in question. The 2\% of trajectories that required
overrides clustered into two classifications: regional trajectories whose
\textit{magnitude} was overly distorted and regional trajectories whose
\textit{shape} was overly distorted.
